```{r echo=FALSE}
library(magrittr)
library(knitr)
library(kableExtra)
options(knitr.kable.NA = "...")
options(kableExtra.html.bsTable = T)
```

# Introduction

## Migration As a Contentious Issue

In the developed world there are few socioeconomic issues as contentious, engaging a significant fraction of the populace over an extended period of time, as migration. There are different types of migrants. Migrants may enter a country to be re-united with their families, to take up employment legally, to seek a better life and work with an illegal status, to seek refuge from famine, war, religious and political persecution. Invariably one type of migrants or the other, become the focus of national and international politics in the developed countries which are the major destinations of migrants. Refugees from Syria and North Africa and their interflow between countries have caused conflicts and fracture within the European Union, and a major factor in Brexit. In the United States, illegal migrants from the south crossing the Mexican border are targets of contentious and partisan politics. In Canada and Australia, opposition to migration never fails to provide fodder for election politics.

National politics aside, migration spawns a myriad of contentious economic, social, ethnic and cultural issues which need to be analyzed and studied to inform the public and the migration policy-makers. Over the decades, there has been a voluminous literature on the study and debate in many countries addressing the following issues. What are migrants’ contribution to the economic development of the host country? Are migrants integrating in the host country? Should government take more initiatives to integrate migrants into the local community? Are migrants occupationally segregated? Do migrants displace the native-born from high-skill jobs or do they take up low-paid jobs unwanted by the native-born? Are professional migrants a significant supplement to the human capital of the host country? Are migrants handicapped in the language of the host country in job search and career advancement? Do migrants take up a disproportionate share of social welfare? Are female migrants participating in the labour market or do they stay home as homemakers? Do migrants benefit from government-subsidized low cost housing at the expense of the native-born? Are migrants residing in enclave communities or trapped in urban ghettos? Do migrants suffer from racial discrimination in the labour market and the housing market? Are migrants upwardly mobile? Do second generation migrants fare better than their parents? Do they fare as well as the native born? Do they have equal educational opportunity? Does migration enrich ethnic and religious diversity or enhance conflict? 

## Censuses and By-censuses

Since our analysis is based on the Hong Kong census, it is pertinent to have a better understanding of census definition of resident population, approach in enumeration and its coverage. Hong Kong census is conducted every ten years in which the entire resident population is enumerated (by a short form) but a fraction of it will be enquired of their detailed characteristics (by the long form). In the mid-point between two censuses, a by-census is conducted. A by-census differs from the census in not having a complete enumeration of the population but focusing only on surveying the detailed characteristics of a fraction of the population by the long form. The size and the characteristics of the entire population in a by-census are inferred from the sample results.

In 1981 one-fifth of the population are surveyed in the census and by-census by the long form. This fraction is reduced to one-seventh in 1986 and further reduced to one-tenth in 2011 as the population grows larger. In the censuses and by-censuses only residents of Hong Kong are surveyed. Over the decades, the approach to enumerating residents changes. In the 1991 census and before, residents present in Hong Kong are enumerated under the de facto approach. All persons who are present in Hong Kong at the reference moment and residents who are temporarily away from Hong Kong are enumerated but only the former are asked detailed questions on demographic and socioeconomic characteristics. In the 1996 census, the de facto approach is changed to the de jure approach. Residents include members of the household usually living in Hong Kong in the six-month period either before or after the reference moment, and those who usually work in Mainland China and Macao. Both groups are enumerated and enquired of their detailed socioeconomic characteristics. Multiple visits are made to interview all members of a household. Information of a household member who is absent for the entire census period may be supplied by other members as far as possible.^[If all members of a household cannot be reached, a self-administered questionnaire which only enquires of their basic characteristics will be left in their mailbox. From 2006 onwards, sampled households can opt to provide information via electronic questionnaires.] From the 2001 census and onward, residents are more finely categorized under the “resident population” approach. The resident population encompasses “Usual Residents” and “Mobile Residents”. “Usual Residents” refer to (1) Hong Kong Permanent Residents who have stayed in Hong Kong for at least three months during the six months before or after the census moment, regardless of whether they are in Hong Kong or not at the census moment; and (2) Hong Kong Non-permanent Residents who are in Hong Kong at the census moment. As for “Mobile Residents”, they are Hong Kong Permanent Residents who have stayed in Hong Kong for at least one month but less than three months during the six months before or after the census moment, regardless of whether they are in Hong Kong or not at the census moment. The change in approach in enumerating Hong Kong residents gives a more precise categorization of Hong Kong Permanent Residents into those residing and working in Hong Kong and those residing in Hong Kong but usually working across the border, as well as Non-permanent Residents who are residing and working in Hong Kong. 

Despite the change in enumeration approach, overall coverage of residents does not seem to change materially across the censuses and should have little effect on our results as our analysis is based on the very large complete dataset of the census. 

## Mainland Chinese Migrants

Mainland Chinese migrants can come to Hong Kong through a number of channels which change over time as policies change in Hong Kong and across the border. In the early years before 1950, the movement of the population across the border between Hong Kong and Guangdong Province was free. In May 1950 the Hong Kong government implemented a quota on the entry of Chinese citizens. With the agreement of the Mainland Chinese authorities, only people granted exit permits (which came to be known as One-way Permits) by the Chinese government were allowed entry into Hong Kong for residence. Subsequent to the implementation of the quota system restricting entry, over time there had been large influxes, at times overwhelming, of migrants who came to Hong Kong illegally. Illegal immigration was stopped by another change in policy abolishing the “reached base” policy in October 1980.^[For a brief history of Mainland Chinese migration to Hong Kong, see Lam and Liu (1998), pp 7-17.] Since then Mainland Chinese migrants can enter Hong Kong only through legal channels including the “One–way Permit” Scheme for family reunion and various professional and talent schemes for employment. The different channels of Mainland Chinese migration to Hong Kong after 1980 and the changing characteristics of the migrants who arrive under these schemes will be the subject of Chapter 3.

In view of the diversity in the background of residents originally from Mainland China who arrive Hong Kong through numerous channels, for the purpose of this book we adopt the simplest definition of a migrant. A migrant is defined by his/her place of birth. Mainland-born residents in Hong Kong in the census are classified as Mainland Chinese migrants, regardless of when they arrive in Hong Kong, whether their entry into Hong Kong is legal or illegal, whether their purpose is for family reunion, work or study, whether their intention is to stay for a short period or long enough to acquire permanent residence status, and whether they enter Hong Kong directly from Mainland China or through another country where they may have stayed a certain length of time and may have acquired citizenship. In short a Mainland Chinese migrant is someone who was born in Mainland China and who came to Hong Kong at a certain stage of his/her life cycle and is surveyed as a resident by the census.

For the purpose of some of the analyses in this book, it is convenient to categorized Mainland Chinese migrants into “old migrants” and “new migrants”. Old migrants are those who have been in Hong Kong for seven or more years. The seven-year duration of stay is a natural criterion for dividing migrants into “old” and “new” because in Hong Kong a residency of seven years qualifies the migrant to apply for permanent residence. The census does not have information on permanent residence but it is believed that most Mainland Chinese migrants do apply and become permanent residents. On the other hand, new migrants are those who have been in Hong Kong less than seven years. There is no information on whether new migrants, especially those who come for education or employment under various schemes, intend to stay beyond seven years and become old migrants with permanent residence status.^[Those who come under the One-way Permit Scheme and some other schemes like Capital Investment Entrant Scheme are probably more clear in their motivation to come to Hong Kong to settle, eventually becoming permanent residents, but less so for those who arrive under the General Employment Policy for Mainland Professionals Residing Overseas, for instance. For details see Chapter 3.] 

```{r chapter1-t1-p, echo=FALSE}
chapter1_t1_p <- read.csv("./prettified/chapter1_t1_p.csv")
colnames(chapter1_t1_p) <- gsub("Year|X", "", colnames(chapter1_t1_p))

kable(chapter1_t1_p, "html",
      caption = "Distribution of Migrant Groups",
      linesep = "",
      booktabs = T,
      align = c(rep("l", 1), rep("c", 8)),
      row.names = F,
      format.args = list(big.mark = ",")) %>%
  column_spec(1, width = "20em", bold = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                font_size = 14, full_width = F) %>%
  pack_rows("All age", 1, 5) %>%
  pack_rows("Age 20-64", 6, 10) %>%
  add_indent(c(3, 4, 8, 9)) %>%
  kableExtra::scroll_box(width = "100%")
```


Table \@ref(tab:chapter1-t1-p) shows the distribution of the Hong Kong born, Mainland migrants (sub-classified into old and new) and others born outside of Mainland China and Hong Kong in number and percentage in the entire population and in the age 20-64 sub-population by census. It should be noted that while the Hong Kong born constitute 58-61% of the population over the period 1981 to 2016, the percentage of Mainland Chinese migrants in the population is rather large, even though it declines from 40% in 1981 to just below 30% in 2016. This percentage is high even comparing to other countries that receive large numbers of migrants. It highlights the importance of studying how well migrants fare in Hong Kong as their economic performance has a significant impact on the well-being of the Hong Kong society.

In Chapter 2, we analyze the demography of Hong Kong over time to show the contribution of migration the population growth in Hong Kong in the context of a long term decline in fertility and the ageing of the population. This chapter places Mainland Chinese migration within the broader demographic background of Hong Kong over time.

Chapter 3 examines the changes in the demographic and socioeconomic characteristics of Mainland migrants over time. In particular, it analyzes in detail changes in Mainland migrants admitted under the One-way Permit scheme after 1997. It also presents other admission schemes based on employment, professional and talent criteria. In summary it shows that cohorts of migrants in recent years are very different in quality and characteristics from the early years. It gives the background information to a better understanding of the cohort effect as an explanation to the changing economic performance of the Mainland migrants.

The economic integration of Mainland migrants as characterized by the convergence or otherwise of their earnings towards those of the Hong Kong born is the subject of Chapter 4. In this chapter we analyze the convergence/ divergence in mean earnings and in earnings distribution of migrants and the Hong Kong born by gender and schooling. The change over time can be explained by the time effect and the cohort effect. We eliminate the cohort by tracking an age cohort over time to show that divergence is followed by convergence in the decade of 2001-2011. We attribute the convergence to a strong time effect arising from the closer economic integration of the economy of Hong Kong and Mainland China.

In Chapter 5 we argue that a key factor to earnings convergence is Putonghua fluency of the Mainland migrants which confers an earnings premium as there is closer economic integration with the Mainland. The Putonghua earnings premium is most pronounced in the finance, services, and the wholesale, retail, restaurant and hotel industries and among professionals and managers. 

Chapter 6 shows that the male labour force participation rate of migrants, both old and new, is similar to the Hong Kong born. While female migrants of child-bearing age tend to have lower participation in the labor market, old migrants beyond child-bearing age have the same labour force participation rate as the Hong Kong born and new migrants actually have a higher participation rate.

In Chapter 7 we compare the proportion of employers, employees and self-employed as well as the industry and occupation distribution of Mainland migrants with the Hong Kong born. There is no evidence of employment segregation of migrants by industries as the migrants and the Hong Kong born have more or less the same employment distribution across industries. There is also no evidence of occupational segregation of university-educated Mainland migrants. However, less educated migrants are segregated to a certain extent into the service and sales occupation and mobility out of these occupations appears low even as they stay longer in Hong Kong.

Housing, in particular public rental housing, accommodation, of Mainland migrants is an issue that attracts public policy debate and legal challenges. Chapter 8 shows that old Mainland migrants are slow in moving out of public housing over time but they are keen to become homeowners through the government subsidized Home Ownership Scheme. The percentage of new migrants residing in public rental housing has dropped significantly to become almost the same as the Hong Kong born as the quality of new arrival cohorts has improved in the recent decade.

In the last chapter, Chapter 9, we address the issue of intergenerational mobility, whether children of migrant parents fare better than their parents in Hong Kong in absolute terms, as well as relative to children of Hong Kong born parents. We show that children of migrant parents, irrespective of whether they were born in the Mainland or in Hong Kong, fare better than their parents in earnings and in educational attainment. They are also more upward mobile across generation than their Hong Kong born counterparts.